Cluster Difference Imaging Photometric Survey (CDIPS) II: Validation of a Jupiter-Sized Planet in IC~2602

### 2020-05-29 paper/ms.tex
update manuscript

\title{Cluster Difference Imaging Photometric Survey (CDIPS) II: A Jupiter-Sized Planet in IC~2602}

@@

\begin{abstract}
We report the validation of the transiting planet TOI 837.01,
a warm Jupiter-sized planet ($R_{\rm p} = 10.XX\,R_\oplus$, $P =
8.3Y\,{\rm d}$), with mass less than X.XX $M_{\rm Jup}$ (3$\sigma$).
TOI$\,$837.01 orbits a V=10.6, T=9.9 G-dwarf in IC~2602 (the
``Southern Pleiades''). It is therefore 50 million years old.
We characterize the system using data from the NASA {\it Transiting
Exoplanet Survey Satellite} (TESS), the ESA Gaia mission,
ground-based photometry, and spectroscopy from CTIO1.5/CHIRON,
XXX/FEROS, and AAT/Veloce.
TOI 837.01 joins the growing ranks of adolescent exoplanets orbiting
bright host stars, and is amenable for studies of planetary
atmospheric and orbital evolution.
\end{abstract}

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.7\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Astrophysical false positive scenarios.}
		{\it Top}: for bound companions,
    {\it Bottom}: for unassociated companions along the same line of
    sight.
		\label{fig:rawzoom}
	}
\end{figure*}

@@

"""Validation is the process of statistically arguing that a
transit-like signal is due to a planet rather than an astrophysical
false positive. """
(CITE blender, Torres+04,05,10; vespa Morton 12, 15; pastis Diaz+14
Santerne+15; triceratops Giacalone \& Dressing 20).
In this study, we provide such statistical arguments.

Exclude possibilities:
"The transits are blended from a background, unassociated eclipsing binary
system or transiting exoplanet: We take the same approach outlined in
Vanderburg et al. (2019). In short, if the transits are a blend from the
background system, the true radius ratio is constrained by the ratio of the
ingress time (T12) to the..."

Given its galactic latitude of only $-6^\circ$, it is not surprising
that the field of \tn\ is crowded.
A number of possible false-positive explanations, including
the usual neighboring, background, and hierarchical eclipsing binary
scenarios, must therefore be considered.

@@

% TODO: there's an even stronger argument advanced by Vanderburg et al
% (2019), and used by Rizzuto+2020.
There's also
\begin{equation}
  \Delta m <
  -\frac{5}{2} \log_{10} (\delta_{\rm obs} t_{12}^2/t_{13}^2).
\end{equation}

@@

Stars brighter than $T=16$, as queried from the Gaia DR2 source
catalog, are shown with orange circles.
The relevant stars for a blend analysis are as follows.
\begin{itemize}
  \item \tn\ = TIC 460205581 (T=9.9). The target star.
  \item Star A = TIC 847769574 (T=14.6). $2.3$'' west. Proper motions
    and parallax imply it is comoving with \tn, though with a physical
    separation of $6.6\pm 0.1\,{\rm pc}$, it may not be bound.
  \item Star B = TIC 460205587 (T=13.1). $5.4$'' north.  This is a
    giant background star.
\end{itemize}
An additional source, TIC 847769581, is $4.9$'' from the target, but
too faint (T=18.8) to be the source of the observed transit signal.

@@

\begin{itemize}
  \item {\it Neighboring blends}.

  \item {\it Hierarchical blends}.
    Limited by i) FEROS + Veloce RVs, and ii) the star being nicely on the
    cluster HR diagram. Any extra hidden close-in binary stars would
    need to be very low-mass M dwarfs, and then the transit duration
		(and lack of chromaticity???) again helps.

  \item {\it Background blends}.
	An argument on the a priori probability of background blends follows from
	counting statistics. Within 360 arcseconds, there are 298 T<16 stars in TIC8
	(i.e., Gaia DR2). At 2.0 arcseconds, SOAR archives dmag(I) of 6. Inside of 2
	arcseconds, we expect $\pi r^2 $ times (Nstar per sq arcsec) = 9e-3 T<16 stars.
	Any with T$\gtrsim$15 outside of 0.3 arcseconds would have been detected by
	SOAR, which brings down the expected number of possible hidden sources by a
	factor of a few.  Within 0.3 arcseconds, where SOAR gets worse than dmag(I) of
	3, we expect 2e-4 T<16 stars. So (order of magnitude), background blends would
	be a ~0.1\% coincidence.

>>>>

\title{Cluster Difference Imaging Photometric Survey (CDIPS) II: Validation of a Jupiter-Sized Planet in IC~2602}

@@

\begin{abstract}
We report the validation of the transiting planet TOI 837.01,
and characterization of the system using data from the NASA {\it Transiting
	Exoplanet Survey Satellite} (TESS), the ESA Gaia mission,
ground-based photometry, and spectroscopy from CTIO1.5/CHIRON,
MPG2.2/FEROS, and AAT/Veloce.
We find that TOI$\,$837 is a $T=9.9$ G-dwarf in IC~2602 (the
``Southern Pleiades''). 
The star is therefore 40$\pm$10 million years old.
The validated planet is warm ($P = 8.YY\,{\rm d}$) and likely
Jupiter-sized. 
However, its transits are grazing: $b > 0.92$ at 3$\sigma$.
From transit photometry alone, the planetary size lies within
0.58--3.87 $R_{\rm Jup}$, due to the degeneracy
between the planetary size and its impact parameter.
From radial velocity monitoring, we find 
that its mass is less than 2.XX $M_{\rm Jup}$ (3$\sigma$).
This mass limit requires the planetary size to be below $\approx$2.X$R_{\rm Jup}$.
Non-planetary explanations for the data, most notably
hierarchical or background eclipsing binaries, are argued to be
highly unlikely on statistical grounds.
TOI 837.01 therefore tentatively joins the ranks of adolescent exoplanets orbiting
bright host stars. Further analysis of its mass, atmosphere, and obliquity could
confirm its planetary nature beyond doubt, and are worthwhile because
young planets do interesting things.
\end{abstract}

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Astrophysical false positive scenarios.}
		{\it Top}: for bound companions,
    {\it Bottom}: for unassociated companions along the same line of
    sight.
		\label{fig:fpscenario}
	}
\end{figure*}

@@

"""Validation is the process of statistically arguing that a
transit-like signal is due to a planet rather than an astrophysical
false positive. """
(CITE blender, Torres+04,05,10; vespa Morton 12, 15; pastis Diaz+14
Santerne+15; triceratops Giacalone \& Dressing 20).

Figure~\ref{fig:fpscenario} provides a visual summary of
the possible astrophysical false positive scenarios.
If the \tn\ signal is not planetary, it must come from
the near-total eclipse of a background star (a BEB), or else a hierarchical companion
of the primary star (an HEB).
It must also be verified that to the angular
resolution possible, the transit signal originates from
\tn, and not a neighboring star (an NEB).

@@

Typically, the shape of the transit enables this argument to be 
scaled to even more restrictive depths (e.g., CITE Vanderburg+2019, Rizzuto+2020, David+2018 1298Tau).
However, since the transits of \tn\ are grazing,
these arguments do not apply.

@@

Stars brighter than $T=16$, as queried from the Gaia DR2 source
catalog, are shown with orange circles.
Given its galactic latitude of only $-6^\circ$, it is not surprising
that the field of \tn\ is crowded.
The relevant stars for a blend analysis are as follows.
\begin{itemize}
  \item \tn\ = TIC 460205581 (T=9.9). The target star.
  \item Star A = TIC 847769574 (T=14.6). $2.3$'' west. Proper motions
    and parallax imply it is comoving with \tn, though with a physical
    separation of $6.6\pm 0.1\,{\rm pc}$, it may not be bound.
  \item Star B = TIC 460205587 (T=13.1). $5.4$'' north.  This is a
    giant background star.
\end{itemize}
An additional source, TIC 847769581, is $4.9$'' from the target, but
too faint (T=18.8) to be the source of the observed transit signal.

@@

\begin{itemize}
  \item {\it Neighboring blends}.
    Ruled out to within 2 arcseconds by Phil Evans.

  \item {\it Hierarchical blends}.
    Limited by i) FEROS + Veloce RVs, and ii) speckle imaging.

  At present our data do not rule out the presence of bound companion
    stars from $\approx$10--100$\,$AU, with masses in the range of
    $\approx$0.1--0.2$\,$M$_\odot$ (Figure~\ref{fig:fpscenario}).
  Roughly 2-3\% of Sun-like stars in the solar neighborhood with
    binary companions have companions in this region of parameter space
  (assuming the Raghavan+2010 CITE log-normal binary
    separation distribution, and a uniform companion mass function.)

  Companion statistics for hierarchical triples are somewhat less 
    robust, but they have been studied, for instance by CITE
    Tokovinin 2014.
  CITE Tokovinin 2014 found that 7\% of Sun-like stars are
    hierarchical triples with two companions in the compact orbit (the
    ``L$_{12}$'' configuration, rather than the ``L$_{11}$''
    configuration of say, $\alpha$~Centauri).
  A priori then, given a Sun-like primary, the probability that it is in a
    hierarchical triple configuration in which the outer orbit
    satisfies the mass and semi-major axis constraints of
    Figure~\ref{fig:fpscenario} is $\approx 7\% \times 3\% = 0.2\%$. 
  This is independent of the issue that the inner pair would require
    finely tuned orbital parameters (i) to eclipse, and (ii) to
    be so compact.

  \item {\it Background blends}.
	An argument on the a priori probability of background blends follows from
	counting statistics. Within 360 arcseconds, there are 298 T<16 stars in TIC8
	(i.e., Gaia DR2). At 2.0 arcseconds, SOAR archives dmag(I) of 6. Inside of 2
	arcseconds, we expect $\pi r^2 $ times (Nstar per sq arcsec) = 9e-3 T<16 stars.
	Any with T$\gtrsim$15 outside of 0.3 arcseconds would have been detected by
	SOAR, which brings down the expected number of possible hidden sources by a
	factor of a few.  Within 0.3 arcseconds, where SOAR gets worse than dmag(I) of
	3, we expect 2e-4 T<16 stars. So (order of magnitude), background blends would
	be a $\sim$0.1\% coincidence.
	%For the background blend to be {\it eclipsing} would require an additional
	%$\sim$2\% coincidence (CITE Kirk+2017 KEBC).
	%Combining the two probabilities yields a prior probability of $\approx$0.002\%.
